Effect of resonant and random excitation on the LHC beam in the context of pulsed hollow electron lens operation.

### 2018-02-17 resex.tex
Added diffusion illustration and caption

\section{Introduction\label{sec:intro}}%force line break with \protect\\
Considering past, current and future high energy and intensity colliders, each new machine has represented a considerable leap in stored beam energy with rising values for future accelerators and colliders (see Table~\ref{tab:stored_energy}). 
\begin{table*}
	\caption{\label{tab:stored_energy}%
		Stored beam energy for different past, present and future colliders. Each new machine represents a leap in stored beam energy.
	}
	\begin{ruledtabular}
		\begin{tabular}{lccccc}
			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}
			& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
			\colrule
			Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
			Number of bunches & 36 & 2220 & 2808 & 2748 & ? \\
			Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
			Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8400 \\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
Recent measurements at the LHC furthermore show that the tails of the transverse beam distribution are overpopulated compared to a Gaussian distribution resulting in a considerable amount of energy being stored in the beam tails alone. In case of the LHC explicitly around 5\% of the beam population is stored in the tails above 3.5~$\sigma$ compared to 0.22\% in case of a Gaussian distribution leading to 19~MJ of stored energy in the tails in case of nominal LHC parameters and 34~MJ in case of HL-LHC~\cite{helreview_valentino}. This leads to the conclusion that a mechanism is needed to deplete the beam tails in a controlled manner \cite{helreview}.

The most obvious idea is to decrease the collimator gaps or scrape the tails with a collimator type device. This approach is however not feasible as it would generate unacceptably large loss spikes due to the high intensity halo population in the LHC. Most promising are methods, which increase the diffusion speed in the region of the halo particles resulting in a smooth and continuous removal of the high amplitude tails while leaving the core of the beam unperturbed. The diffusing halo particles are then intercepted by the collimation system and removed. This concept is also referred to as active halo control compared to a passive system, which only intercepts the halo particles without actively controlling the diffusion speed (see Fig.~\ref{fig:active_halo_control} for illustration).
\begin{figure}[h]
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.7\linewidth]{passive_halo_control.png}
	\end{minipage}
	\begin{minipage}[t]{1.0\linewidth}
		\centering
		\includegraphics[width=0.7\linewidth]{active_halo_control.png}
	\end{minipage}	
	\caption{\label{fig:active_halo_control} Sketch of passive halo control as with the collimation system (top) and active halo control using in addition for example a hollow electron lens (HEL) to control the diffusion speed in the region tail region without affected the beam core (bottom).}
\end{figure}

>>>>

\section{Introduction}
\label{sec:intro}

\begin{table*}
	\caption{\label{tab:stored_energy}%
		Stored beam energy for different past, present and future colliders. Each new machine represents a leap in stored beam energy.
	}
	\begin{ruledtabular}
		\begin{tabular}{lccccc}
			Collider& Tevatron (protons) \cite{tevatron} & LHC 2016 \cite{chamonix2017param}
			& LHC nominal \cite{lhc_design} & HL-LHC \cite{hlcdr} & FCC \cite{fcc_param_2017} \\
			\colrule
			Beam energy [TeV] & 0.98 & 6.5 & 7.0 & 7.0 & 50.0\\
			Number of bunches & 36 & 2220 & 2808 & 2748 & ? \\
			Number of particles per bunch & $2.90\times 10^{11}$ & $1.15\times 10^{11}$ & $1.15\times 10^{11}$ & $2.2\times 10^{11}$ & $1.0\times 10^{11}$\\
			Stored beam energy [MJ] & 1.6 & 265.9 & 362.2 & 678.0 & 8400 \\
		\end{tabular}
	\end{ruledtabular}
\end{table*}

\begin{figure*}
  \begin{minipage}[c]{\columnwidth}
    \includegraphics[width=\columnwidth]{passive_halo_control.png}
    \includegraphics[width=\columnwidth]{active_halo_control.png}
  \end{minipage}
  \hfill
  \begin{minipage}[c]{\columnwidth}
  \includegraphics[width=\columnwidth]{diffusion_illustration}
  \end{minipage}
  \caption{Left: Sketch of passive halo control with a conventional
    collimation system (top) and active halo control with the addition
    of a hollow electron lens (bottom). Right: Illustration of a
    simplified model of active diffusion enhancement in the transverse
    plane. The diffusion coefficient as a function of amplitude
    (orange) is enhanced in a specific amplitude region when the
    hollow beam is turned on (from solid to dashed line). A
    corresponding reduction in beam tail population (black) is created
    (from solid to dashed line).}
  \label{fig:active_halo_control} 
\end{figure*}

Considering past, current and future high energy and intensity
colliders, each new machine has represented a considerable leap in
stored beam energy with rising values for future accelerators and
colliders (see Table~\ref{tab:stored_energy}).

Recent measurements at the LHC furthermore show that the tails of the
transverse beam distribution are overpopulated compared to a Gaussian
distribution resulting in a considerable amount of energy being stored
in the beam tails alone. In case of the LHC explicitly around 5\% of
the beam population is stored in the tails above 3.5~$\sigma$ compared
to 0.22\% in case of a Gaussian distribution leading to 19~MJ of
stored energy in the tails in case of nominal LHC parameters and 34~MJ
in case of HL-LHC~\cite{helreview_valentino}. This leads to the
conclusion that a mechanism is needed to deplete the beam tails in a
controlled manner \cite{helreview}.

The most obvious idea is to decrease the collimator gaps or scrape the
tails with a collimator type device. This approach is however not
feasible as it would generate unacceptably large loss spikes due to
the high intensity halo population in the LHC. Most promising are
methods, which increase the diffusion speed in the region of the halo
particles resulting in a smooth and continuous removal of the high
amplitude tails while leaving the core of the beam unperturbed. The
diffusing halo particles are then intercepted by the collimation
system and removed. This concept is also referred to as active halo
control compared to a passive system, which only intercepts the halo
particles without actively controlling the diffusion speed (see
Fig.~\ref{fig:active_halo_control} for illustration).
